notes

### 2025-07-11 manuscript/reviews/ResponseReview1.Rmd

> 5. Section 2.3.2. I like the idea of making a connection with experimental design,
but am having trouble seeing how the discussion in this section is specific to EHA
and how these comments do not apply also to the “orthodox method”. I think this
section could be improved by providing more specific distinctions between
experimental design implications for EHA vs. the “orthodox method”.

Response: We created a new Figure 1 that makes the intended point, and removed section 2.3.2.

> 6. Line 517: What are model weights? Please explain briefly for the reader,
including how the interpretation differs for loo and waic. 
Also, in the provided
reference (Kruz 2023a), I was only able to find a discussion of dWAIC, rather
than WAIC, so could you please include a definition of or reference for dWAIC in
your context? 
Lastly, when reporting the model weights in the example in the
text, I’d suggest including more than 2 digits so that it is clear if the reported
weights are actually 0 and 1 or if rounding has just made them appear so.

Response: We briefly explain model weights for loo and waic on page 27 of the revised ms.
A discussion of WAIC in Kurz (2023a) can be found in section 4.6.4., so we did not include a reference for dWAIC in the revised ms. Lastly, we print more digits when reporting the model weights.

@@

> 7. Tutorials 3a and 3b: Rather than simply stating that the frequentist models did
not converge and resulted in singular fits, it would be helpful to provide more
discussion. Do you have an idea why the models did not converge? Since this
is a tutorial paper, more explanation is needed. These frequentist examples
don’t add much to the paper in their current form, so I’d either suggest either (a)
providing some suggestions to the reader of alternative approaches or specific
strategies that could be used to define a model that will converge or (b) removing
these examples from the main text and directing the reader to the supplement
instead.

Response: We removed the frequentist examples from the main text and direct the reader to the Tutorials on page 32 of the revised ms.
We mention in Tutorials 3a and 3b that it is common for models to fail to converge in lme4 when they have a reasonably complex random effects structure, and we provide a reference.

@@

> 9. Figure 4: Why is there no bin 40?

Response: Because no responses occurred in the first bin with endpoint 40 ms. However, we added this bin back in Figure 2 (previously 4) of the revised ms to avoid confusion.

> 10. Line 423: The phrase “early responses” is confusing as it implies temporal
ordering of the responses (i.e., early vs. late trials). I’d suggest using “fast
response” or “rapid responses”, or some other phrase along those lines.

Response: We replaced early with fast on page 21 of the revised ms.

@@

> 13. Line 536: To me, “subject-specific” and “marginal” mean the opposite things.
That is, subject-specific is an effect that is conditional on an individual while a
“marginal” effect has been marginalized/averaged over individuals. Could you
please clarify your use of these terms here?

Response: We feel there is no real consensus in the literature about the use of the term marginal. We thus removed this term in the revised ms. But we kept the term in tutorials 2a and 2b where we base our use of the terms on this cited reference: Heiss (2021).

> 14. Lines 584-587: Since this is a tutorial paper, I’d suggest including explanations
for why you make changes (a), (b), and (c).

Response: We now write "The general process is similar to Tutorial 2a, except that (a) we use the person-trial data, (b) we use the symmetric logit link function, and (c) we change the priors (our prior belief is that conditional accuracy values between 0 and 1 are equally likely)." on page 30 of the revised ms. The logit is typically used for logistic regression by experimental psychologists, so we feel that no extra explanation is needed.

@@

> 15. Line 659: What is a “fully varying effects structure”? Please define/explain.

Response: We explain this on pages 22-23 of the revised ms.:
"We also use a "keep it maximal" approach by specifying a full varying (or random) effects structure. This means that wherever possible we include varying intercepts and slopes per participant."

@@

> <b>Organization of the tutorial:</b>The current version of the tutorial is much longer than a typical AMPPStutorial. There are a few possibilities for dealing with this. First, it may be that the tutorial could be split into 2, such as splitting into tutorials 1-3 and tutorial 4 on its own. Another possibility is that the focus could be primarily on the Bayesian models, and then leave the frequentist models for either a supplement, or to only live in the RMarkdown. This possibility may also be beneficial due to the model convergence issues with at least one of the frequentist models (which is common with multilevel data and why many psychological scientists are moving to the Bayesian framework). A third possibility is that the overall tutorial text could be greatly streamlined, with major cuts to both the introduction and the discussion sections.

Response: We focus primarily on the Bayesian models in the revised ms., and keep Tutorial 4.  We also made major cuts to both the introduction and the discussion sections to shorten the ms.

> <b>Clarity and accuracy: </b>
-There are several instances where jargon terms and key concepts pertinent to survival analysis need to be defined and/or described in slightly more detail than what is currently provided. For example, clearer descriptions of censoring and risk sets (plus what it means to be at risk) should be given in the introduction.

Response: We introduce the definition of right-censoring on page 11 of the revised ms. We discuss what it means to be at risk in section A of the Supplemental Material.

@@

> Censoring is a key component of survival analysis that is missing from the introduction

Response: We introduce the concept of right-censoring on page 11 of the revised ms.

> It might be useful to indicate early on the distinctions between multilevel data, single events, and recurring events – and that this particular tutorial is only focused on single events.

Response: We explicitly mention "single event" on page 6 of the revised ms. Also, in section A of the Supplemental Material we mention that this Tutorial does not cover recurrent events, and we visualize the different types of time-to-event data that are obtained in typical RT tasks (detection, discrimination, bistable perception tasks). This should make it clear to the reader that we are dealing with single events that are repeatedly measured within the same individual.

@@

> It wasn’t clear how section 3.1 added value to the tutorial

Response: Section 3 has been removed.

> Section 3.2 – it would be helpful if there was clarification in the text on when it would not be possible to include varying intercepts and slopes (i.e., at least clarifying that random intercepts are a requirement, but sometimes random slopes are removed based on model convergence issues – however, this also begs the question of whether a model should even be presented if it does not converge when it includes random slopes…)

Response: We do not discuss frequentist approaches in the main text anymore (to save space) and only mention that the reader can consult Tutorials 3a and 3b. We agree that a model should not be presented if it does not converge when it includes random slopes. However, to save space, and because this is an active area of research, we do not attempt to clarify when it would not be possible to include varying intercepts and/or slopes. We do mention in Tutorials 3a and 3b that it is common for models to fail to converge in lme4 when they have a reasonably complex random effects structure, and we provide a reference.

@@

> Section 4.1.1. – it would be helpful to clarify that the risk set also pertains to individuals (not just bins)

Response: We clarify this in section A of the Supplemental Material.

> The connection between trials and bins could be made clearer (this is in reference to the text on pg. 18, but I think the clarity would need to come earlier on in the tutorial when describing the data)

Response: This connection should be clear now based on section A of the Supplemental Material, and based on the text on pages 8 and 9.

@@

> Section 4.1.2 was a clearer data description than had been presented up until that point in the tutorial. Again, it is suggested that the manuscript/tutorial could be dramatically reduced in length if information was more concisely but descriptively written in a single place (e.g., data descriptions).

Response: We provide the information on data descriptions in a single place in the revised ms.

> Suggest to be more precise with the language on pg. 21 (line 355) – which data are nested within participants? The trial data? The binned data?

Response: We specify "person-trial data" in the revised ms on page 17.

@@

> pg. 22 – since this is a tutorial, it might also be helpful to unpack in more detail the section about identifying individuals who may be guessing etc – especially since these are very important considerations.

Response: We have no more detail to share on this issue. On page 18 of the revised ms. we now write "In general, it is important to visually inspect the functions first for each participant, in order to identify individuals that may not be following task instructions (e.g., a flat conditional accuracy function at .5 indicates that someone is just guessing), outlying individuals, and/or different groups with qualitatively different behavior.".

> section 4.3.1 – (line 451) it would be helpful if more details is given on how to set the analysis time window. What counts as “enough data”? Given that this is a tutorial, these types of detail are critical for readers to be able to implement the method in their own work

Response: We changed this sentence into: "... because the first few bins
typically contain no responses, one has to select an analysis time window, i.e., a contiguous set of bins for which there is data for each participant." on page 23 of the revised ms.

@@

> section 4.3.1 (line 453) – more detail is needed on the cloglog link function and how it differs from the more commonly known about logit link.

Response: A visual comparison between the cloglog and logit link functions is provided in section D of the Supplemental Material.

> section 4.3.1 (line 456) – what goes into the decision about whether to treat time as a categorical or continuous predictor?

Response: We added the following information on page 23 of the revised ms.:
"Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a categorical or continuous predictor (see also section E of the Supplemental Material). For example, when you want to know if cloglog-hazard is changing linearly or quadraticly over time, you should treat TIME as a continuous predictor.
When you are only interested in the effect of covariates on hazard, you can treat TIME as a categorical predictor (i.e., fit an intercept for each bin), in which case you can choose between reference coding and index coding."

@@

> section 4.3.3. – it might be helpful to give a brief indicator up front to readers about the rationale/purpose of fitting both M0i and M1i, especially given that no interpretation is given for M0i. Similarly, it would be helpful to give a bit more detail on the purpose of the comparison of M0i and M1i (given that this is a didactic tutorial).

Response: We mention now that model M0i is a baseline or reference model, to be compared with model M1i (see page 25 of the revised ms.). We also mention that we only report 2 models in the paper from a larger set of models that are fitted in the file Tutorial_2a.Rmd.

@@

> section 4.3.6. – same thing here – since this is a tutorial, at least brief details on what the purpose of this section is would be helpful. It might also be helpful here to define what is meant by credible interval so that more novice readers don’t confuse it with a confidence interval.

Response: We added the purpose ("To make causal inferences...") and define a credible interval on page 27 of the revised ms.

@@

> pg. 36 – it would be helpful to give more detail on how to interpret the logit-ca scale

Response: We decided not to include more detail because logistic regression is very familiar to experimental psychologists.

> section 4.7 – as previously mentioned, it is possible that tutorial 4 could be a stand alone tutorial. If it’s not though, more details are needed up front to tell readers what the purpose is and why this is useful.

Response: We decided to keep tutorial 4, and discuss its usefulness on pages 40-41 of the revised ms.

### 2025-07-11 manuscript/reviews/ResponseReview1.tex

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{12}
\tightlist
\item
  Line 536: To me, ``subject-specific'' and ``marginal'' mean the
  opposite things. That is, subject-specific is an effect that is
  conditional on an individual while a ``marginal'' effect has been
  marginalized/averaged over individuals. Could you please clarify your
  use of these terms here?
\end{enumerate}
\end{quote}

Response: We feel there is no real consensus in the literature about the
use of the term marginal. We thus removed this term in the revised ms.
But we kept the term in tutorials 2a and 2b where we base our use of the
terms on this cited reference: Heiss (2021).

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{13}
\tightlist
\item
  Lines 584-587: Since this is a tutorial paper, I'd suggest including
  explanations for why you make changes (a), (b), and (c).
\end{enumerate}
\end{quote}

Response: We now write ``The general process is similar to Tutorial 2a,
except that (a) we use the person-trial data, (b) we use the symmetric
logit link function, and (c) we change the priors (our prior belief is
that conditional accuracy values between 0 and 1 are equally likely).''
on page 30 of the revised ms. The logit is typically used for logistic
regression by experimental psychologists, so we feel that no extra
explanation is needed.

@@

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{14}
\tightlist
\item
  Line 659: What is a ``fully varying effects structure''? Please
  define/explain.
\end{enumerate}
\end{quote}

Response: We explain this on pages 22-23 of the revised ms.: ``We also
use a''keep it maximal'' approach by specifying a full varying (or
random) effects structure. This means that wherever possible we include
varying intercepts and slopes per participant.''

@@

\begin{quote}
Censoring is a key component of survival analysis that is missing from
the introduction
\end{quote}

Response: We introduce the concept of right-censoring on page 11 of the
revised ms.

\begin{quote}
It might be useful to indicate early on the distinctions between
multilevel data, single events, and recurring events -- and that this
particular tutorial is only focused on single events.
\end{quote}

Response: We explicitly mention ``single event'' on page 6 of the
revised ms. Also, in section A of the Supplemental Material we mention
that this Tutorial does not cover recurrent events, and we visualize the
different types of time-to-event data that are obtained in typical RT
tasks (detection, discrimination, bistable perception tasks). This
should make it clear to the reader that we are dealing with single
events that are repeatedly measured within the same individual.

@@

\begin{quote}
Section 4.1.2 was a clearer data description than had been presented up
until that point in the tutorial. Again, it is suggested that the
manuscript/tutorial could be dramatically reduced in length if
information was more concisely but descriptively written in a single
place (e.g., data descriptions).
\end{quote}

Response: We provide the information on data descriptions in a single
place in the revised ms.

\begin{quote}
Suggest to be more precise with the language on pg. 21 (line 355) --
which data are nested within participants? The trial data? The binned
data?
\end{quote}

Response: We specify ``person-trial data'' in the revised ms on page 17.

@@

\begin{quote}
pg. 22 -- since this is a tutorial, it might also be helpful to unpack
in more detail the section about identifying individuals who may be
guessing etc -- especially since these are very important
considerations.
\end{quote}

Response: We have no more detail to share on this issue. On page 18 of
the revised ms. we now write ``In general, it is important to visually
inspect the functions first for each participant, in order to identify
individuals that may not be following task instructions (e.g., a flat
conditional accuracy function at .5 indicates that someone is just
guessing), outlying individuals, and/or different groups with
qualitatively different behavior.''.

\begin{quote}
section 4.3.1 -- (line 451) it would be helpful if more details is given
on how to set the analysis time window. What counts as ``enough data''?
Given that this is a tutorial, these types of detail are critical for
readers to be able to implement the method in their own work
\end{quote}

Response: We changed this sentence into: ``\ldots{} because the first
few bins typically contain no responses, one has to select an analysis
time window, i.e., a contiguous set of bins for which there is data for
each participant.'' on page 23 of the revised ms.

@@

\begin{quote}
section 4.3.1 (line 453) -- more detail is needed on the cloglog link
function and how it differs from the more commonly known about logit
link.
\end{quote}

Response: A visual comparison between the cloglog and logit link
functions is provided in section D of the Supplemental Material.

\begin{quote}
section 4.3.1 (line 456) -- what goes into the decision about whether to
treat time as a categorical or continuous predictor?
\end{quote}

Response: We added the following information on page 23 of the revised
ms.: ``Third, one has to choose whether to treat TIME (i.e., the time
bin index t) as a categorical or continuous predictor (see also section
E of the Supplemental Material). For example, when you want to know if
cloglog-hazard is changing linearly or quadraticly over time, you should
treat TIME as a continuous predictor. When you are only interested in
the effect of covariates on hazard, you can treat TIME as a categorical
predictor (i.e., fit an intercept for each bin), in which case you can
choose between reference coding and index coding.''

@@

\begin{quote}
section 4.3.3. -- it might be helpful to give a brief indicator up front
to readers about the rationale/purpose of fitting both M0i and M1i,
especially given that no interpretation is given for M0i. Similarly, it
would be helpful to give a bit more detail on the purpose of the
comparison of M0i and M1i (given that this is a didactic tutorial).
\end{quote}

Response: We mention now that model M0i is a baseline or reference
model, to be compared with model M1i (see page 25 of the revised ms.).
We also mention that we only report 2 models in the paper from a larger
set of models that are fitted in the file Tutorial\_2a.Rmd.

@@

\begin{quote}
section 4.3.6. -- same thing here -- since this is a tutorial, at least
brief details on what the purpose of this section is would be helpful.
It might also be helpful here to define what is meant by credible
interval so that more novice readers don't confuse it with a confidence
interval.
\end{quote}

Response: We added the purpose (``To make causal inferences\ldots{}'')
and define a credible interval on page 27 of the revised ms.

### 2025-07-11 manuscript/Suppl_material/supplementary.Rmd
update ms, notes, and suppl. material

---
title             : no
shorttitle        : "A tutorial on event history analysis"
bibliography      : ["extrareferences2.bib"]
floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
classoption       : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
header-includes:
  - \raggedbottom
---

>>>>

---
title             : no
shorttitle        : "A tutorial on event history analysis"
bibliography      : ["refs_suppl_material.bib"]
floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
classoption       : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
header-includes:
  - \raggedbottom
  - \usepackage{setspace}
  - \captionsetup[figure]{font={stretch=1,normalsize}} 
---

### 2025-07-11 manuscript/notes/notes.Rmd
update ms, notes, and suppl. material

-  Like I said before, the brackets cannot stay like that in Figure 1. It just makes no sense. Instead, just use words and no brackets. e.g., The first bin is for responses up to 100ms (or whatever) and the last bin is for responses between 901 and 1000. Or just do [1,100] and [901,1000], if that's easier.

@@

>>>>

-  Like I said before, the brackets cannot stay like that in Figure 1. It just makes no sense. Instead, just use words and no brackets. e.g., The first bin is for responses up to 100ms (or whatever) and the last bin is for responses between 901 and 1000. Or just do ]1,100] and [901,1000], if that's easier.

-> I use words now, and keep 1 example for illustration.

@@

-> I increased the size of both figures and made the figure (and table) captions smaller using
header-includes:
  - \usepackage{setspace}
  - \captionsetup[figure]{font={stretch=1,normalsize}}

This also decreased Table sizes, and footnote size. I think this is actually fine.
